Ultra-deep $K_s$-band Imaging of the \textit{Hubble} Frontier Fields

### 2016-02-22 Paper0/ms.tex
further fix text for latest analysis

We present ultra-deep $K_s$-band imaging of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, MACS-0717 and MACS-1149 (\xxx\ include MOSFIRE???), which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 4.5$\,\mum.  \textit{VLT}/HAWKI integrations of the first two fields reach 5-$\sigma$ limiting depths of $K_s=26.3$ (AB) and have seeing FWHM$\sim0\farcs4$.  Shorter MOSFIRE integrations of the latter two northern fields reach limiting depths $K_s=$\xxx\ with seeing FWHM=\xxx.  In all cases the $K_s$-band mosaics cover the primary cluster and parallel \textit{HST}/ACS+WFC3 fields, and the total area of the $K_s$-band coverage is \xxx\ arcmin$^2$.  The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$) and the IRAC 3.6\,\mum\ passband.  While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide significant constraints on both the redshifts and the stellar-population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).  Reduced, aligned mosaics of the four survey fields are provided accompanying this manuscript.  Time has been awarded for similar HAWKI coverage of the final two Frontier Fields Abell 370 and Abell S1063, with observations expected by the end of 2015.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%  Summary table
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{deluxetable*}{ccccrccc} 
\tabletypesize{\footnotesize}
\tablewidth{0pt} 
\tablecaption{Field summary \label{tab:summary}} 
\tablehead{ 
\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth$^\mathrm{a}$} & \colhead{FWHM}}
\startdata 
Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22 & 25.5 h & 25.9 & $0\farcs38$ \\ 
Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & 25.5 h & 25.9 & $0\farcs38$ \\ 
MACS-0717    & $07:17:34.0$ & $+37:44:49$ & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27, 2016 Jan 21--22  & 0/0 h\footnotemark[2] & 25.0/25.0 & $0\farcs43$/$0\farcs00$ \\ 
MACS-1149    & $11:49:36.3$ & $+22:23:58$ & \textit{Keck}/MOSFIRE  & 2015 Feb 24, 2016 Jan 21--22      & \xxx\footnotemark[2] & \xxx & \xxx \\
''           & ''           & ''          & \textit{VLT}/HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & \xxx & \xxx & \xxx
\enddata 
%\vspace{-0.8cm} 
\footnotetext[1]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
\footnotetext[2]{The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster/parallel \textit{HST} fields, which have the different characteristics as indicated, respectively.}
\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
\end{deluxetable*}
%\vspace{1cm}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   2MASS comparison
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!t]
\epsscale{1.18}
\plotone{2mass_phot.pdf} % Figures/exposure_map.pdf
\caption{Comparison of bright-star photometry to the 2MASS public catalog.  The Abell 2744 and MACS 0416 fields at high galactic latitude have few stars for cross calibration, and the photometric calibration is determined from observations of standard stars on nights with photometric transparency conditions.  At relatively low galactic latitude ($b=+21$), the MACS 0717 field has many bright stars, and the photometric calibration in that field is defined from the 2MASS comparison (hence, $\Delta_\mathrm{2mass}\equiv0$, as indicated for that field).  We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}  
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Image quality figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht]
\epsscale{1.18}
\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.  The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ and $\beta\sim2$, which have substantially more flux at radii $r>\mathrm{FWHM}$ than Gaussian profiles with the same half-widths (orange curves).  These extended profiles appear to be characteristic of deep $K_s$-band images; the stellar profiles from the HUGS GOODS-South D2 HAWK-I $K_s$ image \citep{hugs} (lower right panel), created by a different team with a different reduction pipeline, are nearly identical to those of the Frontier Fields mosaics presented here. \xxx\ I only show the parallel field region of MACS0717, which has significantly better image quality from the Jan. run.  You can see a few of the broader profiles in the figure; there are many more of these if you include the whole cluster area.  \label{fig:cog}}  
\end{figure*}

@@

A crucial innovation of the image processing of the present analysis compared to previous works is that we combine all of the individual background-subtracted raw images (e.g., \xxx\ 4000 files for 1000 exposures times 4 detectors for the \mbox{Abell 2744} field) into the final output mosaic with the ``drizzle'' algorithm (\citealp{fruchter:02}, as implemented in the DrizzlePac/AstroDrizzle package; \citealp{drizzlepac}).  The benefits of the drizzle algorithm  preserving robust noise properties of the output mosaics will be discussed in detail below in \S\xxx.  Here we simply indicate that the drizzle implementation provides additional benefits in 1) trivial application of relative weights of the input images in creating the final mosaics, 2) easily applying the astrometric alignment and geometric distortion stored in the FITS headers, and 3) allowing the definition of an arbitrary output pixel grid without any additional image resampling.

Final mosaics of each field with $0\farcs1$ pixels are drizzled from the input exposures, $i$, with weights
\[
w^{-1} = F^2\cdot B \cdot \mathrm{FWHM}^\alpha / \sqrt{1-e^2}, 
\]
where $F$ is the factor to scale the exposure to the photometric system ($F\equiv1$ for the OB taken under photometric conditions, $F\gtrsim1$ otherwise), $B$ is the measured background, FWHM is the full width at half maximum of stars identified in the field, and $e$ is the ellipticity of the stellar point spread function (PSF).  The parameter $\alpha$ allows for optimizing the image quality of the final mosaic.  Increasing $\alpha$ puts larger weight on the exposures with the best image-quality while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.

Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies. The mosaics are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.  Inverse variance maps are shown in Appendix \xxx\.  The final image and inverse-variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).  The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity for faint galaxies.  In the sections below we describe the characteristics of the mosaics in more detail.

@@

Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.  The image quality of the $K_s$-band mosaics is excellent, with stellar FWHM $0\farcs40$ and $0\farcs38$ in the Abell 2744 and MACS 0416 fields, respectively, thanks, in large part, to the execution of the HAWK-I observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.  Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.  Also shown in Fig.~\ref{fig:cog} are stellar curves of growth measured from the HAWK-I UDS and GOODS Survey \citep[HUGS;][]{hugs}, which are nearly identical in both FWHM and profile shape.  We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles. \xxx\ not clear what HAWK-I ETC assumes, the description just says ``FWHM of the seeing disk'' \xxx.  Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging than the redder \textit{Spitzer} IRAC bands.

\xxx\ add description of MOSFIRE clusters.  IQ not uniform across the MACS0717 field.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Depth figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\epsscale{1.18}
\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
\caption{\textit{Top panel:} \xxx Noise in empty apertures as a function of aperture diameter, normalized to the smallest $D=0\farcs2$ aperture.  The dashed and dotted lines show the expectations for uncorrelated and perfectly-correlated noise within the aperture.  \textit{Bottom panel:} The product of the inverse curves-of-growth (Fig.\ref{fig:cog} and the empty aperture noise gives the point-source sensitivity as a function of aperture size.  The S/N is maximized with an aperture just larger than the FWHM.  For reference, the colored dashed lines in the bottom panel indicate the depth computed assuming uncorrelated noise within the apertures ($\sigma \propto D$) and the per-pixel variance reported in the weight maps.)
\label{fig:depth}}  
\end{figure}
%%%%%%

@@

The fact that the drizzle-combined images in Fig.~\ref{fig:drizzle_cutout} (panels \textit{a} and \textit{b}) appear to the eye to be noisier than their SWarped counterparts (panels \textit{c} and \textit{d}) is simply the result of a smoothing of the pixel noise in the latter case rather than reflecting true underlying differences in their pixel-to-pixel variance.  \cite{casertano:00} discuss how the noise properties of correlated output pixels are affected by the relative sizes of the input and output pixel grids and the adopted drizzle parameters.  With the drizzle algorithm and a small pixfrac, a given input pixel will map to only a single output pixel and therefore correlations between the output pixels should be minimal\footnote{Inter-pixel capacitance and other forms of crosstalk intrinsic to the IR detectors effectively place a lower limit on the correlations between adjacent pixels \citep{hilbert:ipc}.}.

@@

In terms of understanding the noise properties of the images, a key point here is not only that depression at small separations indicates the presence of correlations between adjacent pixels but also the fact that correlations cause the apparent overall r.m.s. to be decreased at all separations.  That the pixel variances can't be trusted in the presence of such correlations is the primary reason why significant effort has been devoted to placing random ``empty apertures'' across images in order to characterize their noise properties \citep[e.g., ]{quadri:musyc, whitaker:nmbs, skelton:14}.  Now by eliminating the pixel correlations with the drizzle combination (with a small pixfrac) we have shown that the pixel statistics are robust and it is now trivial to compute the expected variance within an arbitrary photometric aperture, for example:
\begin{equation}
\sigma_D^2 = \sigma_\mathrm{pix}^2\cdot \pi/4 \cdot D^2,
\label{eq:aper_var}
\end{equation}
for circular apertures with diameter, $D$, in pixels, and where $\sigma_\mathrm{pix}^2$ is the per-pixel variance determined from an analysis such as that in Fig.~\ref{fig:corr_noise}.

The product of Eq.~\ref{eq:aper_var} and the inverse of the curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.  For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs6$, just larger than the FWHM of point-source profiles.  The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are \xxx\ 25.8, 25.9, and 25.0 in the Abell 2744, MACS-0416, and MACS-0717 fields, respectively.  The HAWK-I images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.

We conclude here with a brief comment on the ``HUGS'' ultra-deep HAWK-I images of the CANDELS survey fields.  \cite{hugs} report a final depth of 26.5 AB for point sources within $D=0\farcs4$ apertures for their deepest field, ``GOODS-D1'', which has an exposure time of 31.5 hours similar to the deep HAWK-I fields summarized in Table~\ref{tab:summary}.  \cite{hugs} assume ideal noise properties that likely suffer some residual correlations between adjacent pixels such as that indicated by the red curves in Fig.~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.  This effect ($\sim$0.4 mag) along with the relative exposure times ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields HAWK-I images described here.  The comparison must be made because any additional half magnitude increase depth is exponentially more expensive to obtain!

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Cutout figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\epsscale{1.18}
\plotone{image_cutouts.pdf} % Figures/exposure_map.pdf
\caption{Image cutouts of the MACS 0416 parallel field.  The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $10\arcsec$ \xxx\ [check] on a side.
\label{fig:cutout}}  
\end{figure}

>>>>

We present ultra-deep $K_s$-band imaging of all six of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, Abell S1063, Abell 370, MACS-0717 and MACS-1149, which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 1.6$\,\mum and 3.6--4.5\,\mum, respectively.  \textit{VLT}/HAWKI integrations of the first four fields reach 5-$\sigma$ limiting depths of $K_s\sim26.0$ (AB, point sources) and have seeing FWHM$\sim0\farcs4$.  Shorter MOSFIRE integrations of the MACS-0717 (MACS-1149) field better observable in the north reach limiting depths $K_s$=25.5 (25.1) with seeing FWHM$\sim$$0\farcs4$ ($0\farcs5$).  In all cases the $K_s$-band mosaics cover the primary cluster and parallel \textit{HST}/ACS+WFC3 fields, and the total area of the $K_s$-band coverage is \xxx\ arcmin$^2$.  The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$) and the IRAC 3.6\,\mum\ passband.  While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide significant constraints on both the redshifts and the stellar population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).  Reduced, aligned mosaics of all six survey fields are provided accompanying this manuscript.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%  Summary table
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{deluxetable*}{ccccrccc} 
\tabletypesize{\footnotesize}
\tablewidth{0pt} 
\tablecaption{Field summary \label{tab:summary}} 
\tablehead{ 
\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth$^\mathrm{a}$} & \colhead{FWHM}}
\startdata 
Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22       & \xxx & \xxx & \xxx \\ 
Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & \xxx & \xxx & \xxx \\ 
MACS-0717    & $07:17:34.0$ & $+37:44:49$ & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27, 2016 Jan 21--22  & 0/0 h\footnotemark[2] & 25.0/25.0 & $0\farcs43$/$0\farcs00$ \\ 
MACS-1149    & $11:49:36.3$ & $+22:23:58$ & \textit{Keck}/MOSFIRE  & 2015 Feb 24, 2016 Jan 21--22      & \xxx\footnotemark[2] & \xxx & \xxx \\
''           & ''           & ''          & \textit{VLT}/HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & \xxx & \xxx & \xxx
\enddata 
%\vspace{-0.8cm} 
\footnotetext[1]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
\footnotetext[2]{The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster/parallel \textit{HST} fields, which have the different characteristics as indicated, respectively.}
\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
\end{deluxetable*}
%\vspace{1cm}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   2MASS comparison
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!t]
\epsscale{1.18}
\plotone{2mass_phot.pdf} % Figures/exposure_map.pdf
\caption{Comparison of bright-star photometry to the 2MASS public catalog.  The photometric calibration of the $K_s$-band images were determined from standard star observations taken commensurate with the science exposures.  The uncertainties shown are taken from the 2MASS catalog; these stars are measured at very high S/N in the deep $K_s$-band mosaics.  We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}  
\end{figure}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Drizzle figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!t]
\epsscale{1.18}
\plotone{ivar_mosaics.pdf} % Figures/exposure_map.pdf
\caption{Inverse variance maps of the $K_s$-band mosaics.  The orientation of the fields is the same as that shown in Fig.~\ref{fig:layout}.  The four HAWK-I chips and two MOSFIRE pointings are clearly visible for the first four and last two fields, respectively.
\label{fig:ivar_mosaics}}  
\end{figure}

A crucial innovation of the image processing of the present analysis compared to previous works is that we combine all of the individual background-subtracted raw images (e.g., \xxx\ 4000 files for 1000 exposures times 4 detectors for the \mbox{Abell 2744} field) into the final output mosaic with the ``drizzle'' algorithm (\citealp{fruchter:02}, as implemented in the DrizzlePac/AstroDrizzle package; \citealp{drizzlepac}).  The benefits of the drizzle algorithm  preserving robust noise properties of the output mosaics will be discussed in detail below in \S\ref{s:depth}.  Here we simply indicate that the drizzle implementation provides additional benefits in 1) trivial application of relative weights of the input images in creating the final mosaics, 2) providing infrastructure to apply the astrometric alignment and geometric distortion stored in the FITS headers, and 3) allowing the definition of an arbitrary output pixel grid without any additional image resampling.

Final mosaics of each field with $0\farcs1$ pixels are drizzled from the input exposures with weights
\begin{equation}
w^{-1} = F^2\cdot B \cdot \mathrm{FWHM}^\alpha / \sqrt{1-e^2}, 
\label{eq:weight}
\end{equation}
where $F$ is the factor to scale the exposure to the photometric system ($F\equiv1$ for the average of exposures in the OB taken under photometric conditions, $F\gtrsim1$ otherwise), $B$ is the measured background, FWHM is the full width at half maximum of stars identified in the field, and $e$ is the ellipticity of the stellar point spread function (PSF).  The parameter $\alpha$ allows for optimizing the image quality of the final mosaic.  Increasing $\alpha$ puts larger weight on the exposures with the best image-quality while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.

Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies. The mosaics are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.  We also compute the robust NMAD scatter \citep{brammer:08} of empty sky pixels within the same cells to empirically calibrate the inverse variance maps that are shown in Fig.~\ref{fig:ivar_mosaics} (see also \S\ref{s:depth}).  The final image and inverse variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).  The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity for faint galaxies.  In the sections below we describe the characteristics of the mosaics in more detail.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Image quality figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht]
\epsscale{1.18}
\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.  The inverse aperture corrections for point sources in $D=0\farcs6$ apertures are indicated in each panel.  The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ in most cases and $\beta\sim2.1$.  These extended Moffat profiles appear to be characteristic of deep $K_s$-band images and they have substantially more flux at large radii ($r>\mathrm{FWHM}/2$) than Gaussian profiles with the same FWHM (orange curves in the left panels).  \label{fig:cog}}  
\end{figure*}

@@

Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.  The image quality of the $K_s$-band mosaics is excellent with stellar FWHM$\lesssim0\farcs4$ for the deep HAWK-I fields, thanks in large part to the execution of the HAWK-I observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.  The classical scheduling of the MOSFIRE observations does not allow such control over the seeing conditions.  The result is that the image quality of the MACS-0717 and MACS-1149 fields is somewhat degraded in comparison, at $0\farcs42$--$0\farcs54$, and the image quality differs measurably between the two MOSFIRE pointings in the MACS-0717 field.  Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.  We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.  Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging and the typical apparent size of distant galaxies \citep[median $r_e\sim0\farcs1$--$0\farcs2$ at $z>4$;][]{shibuya:15} than the redder \textit{Spitzer} IRAC bands (cf. FWHM$\sim$$1\farcs7$--$2^{\prime\prime}$).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Depth figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\epsscale{1.18}
\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
\caption{\textit{Top panel:} The product of the inverse curves-of-growth (Fig.~\ref{fig:cog}) and the predicted variance within a photometric aperture (Eq.\ref{eq:aper_var}) gives the point-source sensitivity as a function of aperture size.  The S/N is maximized with an aperture somewhat larger than the FWHM; depths evaluated at $D=0\farcs6$ near the maxima are indicated.)
\label{fig:depth}}  
\end{figure}
%%%%%%

@@

The fact that the drizzle-combined images in Fig.~\ref{fig:drizzle_cutout} (panels \textit{a} and \textit{b}) appear to the eye to be noisier than their SWarped counterparts (panels \textit{c} and \textit{d}) is simply the result of a smoothing of the pixel noise in the latter case rather than reflecting true underlying differences in their pixel-to-pixel variance.  \cite{casertano:00} discuss how the noise properties of correlated output pixels are affected by the relative sizes of the input and output pixel grids and the adopted drizzle parameters.  With the drizzle algorithm and a small pixfrac, a given input pixel will map to only a single output pixel and therefore correlations between the output pixels should be minimal\footnote{Inter-pixel capacitance and other forms of crosstalk intrinsic to the IR detectors place a lower limit on the correlations between adjacent pixels \citep{finger:05, hilbert:ipc}.}.

@@

In terms of understanding the noise properties of the images, a key point here is not only that depression at small separations indicates the presence of correlations between adjacent pixels but also the fact that correlations cause the apparent overall r.m.s. to be decreased at all separations.  That the pixel variances can't be used directly in the presence of such correlations \citep{casertano:00} is the primary reason why significant effort has been devoted to placing random ``empty apertures'' across images in order to characterize their noise properties \citep[e.g.,][]{quadri:musyc, whitaker:nmbs, skelton:14}.  Now by eliminating the pixel correlations with the drizzle image combination technique (with a small pixfrac) we have shown that the pixel statistics are robust and it is now trivial to compute the expected variance within an arbitrary photometric aperture, for example:
\begin{equation}
\sigma_{\mathrm{aper}, D}^2 = \sigma_\mathrm{pix}^2\cdot \pi/4 \cdot D^2,
\label{eq:aper_var}
\end{equation}
for circular apertures with diameter, $D$, in pixels, and where $\sigma_\mathrm{pix}^2$ is the per-pixel variance determined from an analysis such as that in Fig.~\ref{fig:corr_noise}.

The product of Eq.~\ref{eq:aper_var} and the inverse of the curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.  For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs6$, just larger than the FWHM of point-source profiles.  The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are indicated in Fig.~\ref{fig:depth} and listed in Table~\ref{tab:summary}.  Reaching 26th magnitude (AB), the HAWK-I images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.

We conclude here with a brief comment on the ``HUGS'' ultra-deep HAWK-I images of the CANDELS GOODS-S and UDS survey fields.  \cite{hugs} report a final depth of 26.5 AB for point sources within $D=0\farcs4$ apertures for their deepest field, ``GOODS-D1'', which has an exposure time of 31.5 hours similar to the deep HAWK-I fields summarized in Table~\ref{tab:summary}.  \cite{hugs} assume ideal noise properties that likely suffer some residual correlation between adjacent pixels such as that indicated by the red curves in Fig.~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.  This effect ($\sim$0.4 mag) along with the relative exposure time difference ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields HAWK-I images described here.  The comparison must be made because any additional half magnitude increase depth is exponentially more expensive to obtain!

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Cutout figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\epsscale{1.18}
\plotone{image_cutouts.pdf} % Figures/exposure_map.pdf
\caption{Image cutouts of the MACS 0416 parallel field.  The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $10\arcsec$ \xxx\ \textcolor{red}{check, different now} on a side.
\label{fig:cutout}}  
\end{figure}
